```{r setup}
knit_hooks$set(webgl = hook_webgl)
```

```{r, echo = FALSE}
source("~/work/ecodatascripts/vars/heights/prep.R")  
source("~/work/ecodatascripts/vars/heights/canopy/load_canopy.R")
source("~/work/ecodatascripts/vars/heights/vis_functions.R")
library(dplyr, warn.conflicts=FALSE, quietly=TRUE)
library(magrittr)
library(ggplot2)
library(rgl)
library(bbmle)

fits <- readRDS("../temp/fits.rds")
stats <- readRDS("../temp/stats.rds")
becos <- stats[stats$Species == "beco", ]
abbas <- stats[stats$Species == "abba", ]
```

## HH fits
Gompertz:
$H = \beta*e^{log(\frac{\gamma}{\beta})e^{-\alpha DBH}}$

Power:
$H = \gamma + \beta DBH^{\alpha}$

AIC values for gompertz and power fits.
```{r}
abbas %>%
  arrange(Year, AICc)
becos %>%
  arrange(Year, AICc)
```


# Gompertz models with canopy height only (the current version used for height prediction).
## BECO fits:
```{r, webgl=TRUE, cache=TRUE, echo=FALSE}
spec <- "beco"
model <- "gompertz"
inds <- "can"
hh <- TRUE
years <- c(99, 11)
preds <- get_preds(spec, years, modtype = model, inds = inds, hh=hh)
plot_preds(preds)
add_pred_lines(preds)
add_observed(preds)
```

## ABBA fits:
```{r, webgl=TRUE, cache=TRUE, echo=FALSE}
spec <- "abba"
model <- "gompertz"
inds <- "can"
hh <- TRUE
years <- c(99, 11)
preds <- get_preds(spec, years, modtype = model, inds = inds, hh=hh)
plot_preds(preds)
add_pred_lines(preds)
add_observed(preds)
```

## Best fits
Gompertz full model is fitting best for ABBA.
```{r, echo=FALSE}
ind <- grep("gompertz_abba_[[:digit:]]+_full", names(fits))
f99 <- fits[[ind[1]]]
f11 <- fits[[ind[2]]]
```
### 99
```{r}
summary(f99)
```
### 11
```{r}
summary(f11)
```


# Profiles
## ABBA, gompertz full
```{r, cache=TRUE, echo=FALSE}
can_func = "can_hh"
spec = "ABBA"
method = "Nelder-Mead"
source("~/work/hh/gompertz/full/model.R")

## 99
fit <- f99
yr <- 99
p1 <- profile(fit, which = names(ps)[!(names(ps) %in% c("sd", "b", "b2", "a", "a2"))])  # can't profile 'b' (flat)
# ps <- coef(p1)
# names(ps) <- gsub("\\..*", "", names(ps))
# dat <- prep_hh(dat=tp, yr=yr, spec=spec, can_func=can_func)
# fit <- run_fit(dat, ps, yr)
plot(p1)
```

## BECO, power/gompertz canopy only (11) and full gompertz (99)
```{r, echo=FALSE}
ind <- grep("power_beco_[[:digit:]]+_can", names(fits))
f11 <- fits[[ind[2]]]

ind <- grep("gompertz_beco_[[:digit:]]+_full", names(fits))
f99 <- fits[[ind[1]]]

can_func = "can_hh"
spec = "BECO"
method = "Nelder-Mead"
```

### 99
```{r}
summary(f99)
```

```{r, echo=FALSE}
# source("~/work/hh/gompertz/full/model.R")
# summary(f99)
# fit <- f99
# yr <- 99
# 
# p1 <- profile(fit)
# # p1 <- profile(fit, which = names(ps)[!(names(ps) %in% c("sd", "b", "b2", "a", "a2"))])  # can't profile 'b' (flat)
# ps <- coef(p1)
# names(ps) <- gsub("\\..*", "", names(ps))
# dat <- prep_hh(dat=tp, yr=yr, spec=spec, can_func=can_func)
# fit <- run_fit(dat, ps, yr)
# 
# plot(p1)
```

### 11
```{r, cache=TRUE}
source("~/work/hh/power/can/model.R")
summary(f11)
p1 <- profile(fit)
plot(p1)
```
