```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
                      echo=FALSE, warning=FALSE, message=FALSE)
```

```{r setup, results='hide'}
source("setup.R")
source("functions.R")

## Quantiles
qDBH <- dat %>% group_by(YEAR) %>% do(qFunc(., "DBH"))  # DBH quantiles by year
qHT <- dat %>% group_by(YEAR) %>% do(qFunc(., "HT"))    # HT quantiles by year

```

## Linear models for HT vs. DBH by size classes
* 4 roughly equal size classes (quantiles), created using both years combined.
So, 99 has fewer trees in large size class and more in small classes.

```{r samp_size}

numSizes <- 4

## Set quantile bounds for size class classification
probs <- c(0, 0.25, 0.5, 0.75, 1.0)
dat <- classify(dat, probs)
t1 <- table(dat[, c("DBHCL","HTCL","YEAR")])
addmargins(t1)

```

```{r models}

## All interactions
m1 <- lm(HT ~ DBH * I(DBHCL) * I(YEAR) * CANHT * ELEV, data = dat)
sm1 <- summary(m1)
a1 <- anova(m1)

## No interactions
m2 <- lm(HT ~ DBH + I(DBHCL) + I(YEAR) + CANHT, data = dat)
sm2 <- summary(m2)
a2 <- anova(m2)

## Interactions b/w DBH and CANHT
m3 <- lm(HT ~ DBH * CANHT + I(YEAR) + I(DBHCL), data = dat)
sm3 <- summary(m3)
a3 <- anova(m3)

## Interactions => DBH:CANHT, DBH:YEAR
m4 <- lm(HT ~ DBH * CANHT + I(YEAR) + I(DBHCL) + DBH:I(YEAR), data = dat)
sm4 <- summary(m4)
a4 <- anova(m4)

## Interactions => DBH:CANHT, DBH:YEAR, CANHT:YEAR
m5 <- lm(HT ~ DBH * CANHT + I(YEAR) + I(DBHCL) + DBH:I(YEAR) + CANHT:I(YEAR), data = dat)
sm5 <- summary(m5)
a5 <- anova(m5)

m6 <- lm(HT ~ DBH + I(DBHCL) + ELEV + CANHT + DBH:CANHT, data = dat)
sm6 <- summary(m6)
a6 <- anova(m6)

m7 <- lm(HT ~ DBH + I(DBHCL) + ELEV  + DBH:CANHT, data = dat[dat$YEAR==99,])
m8 <- lm(HT ~ DBH + I(DBHCL) + ELEV + CANHT + DBH:CANHT, data = dat[dat$YEAR==99,])

m9 <- lm(HT ~ DBH + I(DBHCL) + ELEV + CANHT + DBH:CANHT, data = dat[dat$YEAR==11,])
m10 <- lm(HT ~ DBH + I(DBHCL) + ELEV + DBH:CANHT, data = dat[dat$YEAR==11,])

```

- Model with year and DBHCL as indicator variables, canopy height, DBH, ELEV as continuous.  All interactions are also included.
```{r}
## sm1
```

- Model with no interactions, no ELEV
```{r}
## sm2
```

- Only interactions between DBH and canopy height, no ELEV
```{r}
## sm3
```

## Visualize
**Look at them separately by DBH size classes**
### Model 1: all interactions, continuous variables: {ELEV, CANHT, DBH}, factors: {DBHCL, YEAR}

#### Separate panels for DBH classes
Any remaining nonlinearites in the lines are from elevation effects (ie. the squiggles).

```{r m1_visuals}

## dat$pred <- predict(m6, newdata = dat)
## chts <- unlist(quantile(dat$CANHT, na.rm=T))
## chts[1] <- 0

## for (i in 1:numSizes) {
##     for (j in 1:(length(chts)-1)) {
##         p <- ggplot(na.omit(dat[dat$CANHT > chts[[j]] & dat$CANHT <= chts[[j+1]] & dat$DBHCL == i,]),
##                     aes(DBH, HT, color=YEAR)) + geom_point(alpha=0.5) +
##                         geom_line(aes(DBH, pred, color=YEAR), lwd=1.2) +
##                             facet_wrap(~ CANHT) +
##                                 ggtitle(paste0("HT vs DBH (DBHCL == ", j, " , split by Canopy Height"))
##         print( p )
##     }
## }

```


## Fit from 99

```{r year_model}
dat$pred99 <- predict(m7, newdata = dat)
dat$pred11 <- predict(m10, newdata = dat)

chts <- unlist(quantile(dat$CANHT, na.rm=T))
chts[1] <- 0


## for (i in 1:numSizes) {
##     for (j in 1:(length(chts)-1)) {
##         p <- ggplot(na.omit(dat[dat$CANHT > chts[[j]] & dat$CANHT <= chts[[j+1]] & dat$DBHCL == i,]),
##                     aes(DBH, HT, color=YEAR)) + geom_point(alpha=0.5) +
##                         geom_line(aes(DBH, pred, color=YEAR), lwd=1.2) +
##                             facet_wrap(~ CANHT) +
##                                 ggtitle(paste0("HT vs DBH (DBHCL == ", j, " , split by Canopy Height"))
##         print( p )
##     }
## }

trans <- levels(dat$TRAN)
for (tran in trans) {

    if (nrow(dat[dat$TRAN == tran,]) > 0) {

        for (i in 1:numSizes) {
            dd <- dat[dat$DBHCL == i & dat$TRAN == tran,]
            p99 <- dd[dd$YEAR == 99, ]
            p11 <- dd[dd$YEAR == 11, ]
            
            p <- ggplot(na.omit(dd), aes(DBH, HT, color=YEAR)) + geom_point(alpha=0.5) +
                geom_line(data=na.omit(p99), aes(DBH, pred99), color = "blue", lty = 2) +
                    geom_line(data=na.omit(p11), aes(DBH, pred99), color="red", lty=2) +
                        geom_line(data=na.omit(p99), aes(DBH, pred11), color="blue") +
                            geom_line(data=na.omit(p11), aes(DBH, pred11), color="red") +
                                facet_wrap(~ TPLOT) +
                                    ggtitle(paste0("Transect: ", tran, ", DBHCL == ", i))
            print( p )
        }
    }
}

```
