---
title: "HH Nonlinear Fits"
date: "Last modified: `r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    number_sections: false
    theme: spacelab
    highlight: zenburn
---

```{r global_opts, include=FALSE}
knitr::opts_chunk$set(fig.path='Figs/', fig.width=12, fig.height=8,
                      echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE)
```

```{r setup}

source("./setup.R")
source("./functions.R")
abbas <- dat[dat$SPEC == "ABBA", ]
becos <- dat[dat$SPEC == "BECO", ]

```

# Info
[code](http://github.com/ghandi9000/hh) repository

- Modeled for 99, and 11.
- Boths models predicted for both years.
- Issues: a couple of the BECO fits are bad from 2011 and need to be rerun.

---------------

# Full Gompertz

__Full Model__ {dbh, canopy, elevation}

$$
HT = \beta e^{\log(\frac{\gamma}{\beta})e^{-\alpha DBH}}
$$

where $\gamma$ is DBH height,

$\alpha = a + a_1 ELEV + a_2 CANHT + a_3 ELEV*CANHT$,

$\beta = b + b_1 ELEV + b_2 CANHT + b_3ELEV*CANHT$

--------------------------------------------------

## ABBA

ABBA **1999** estimates
```{r gomp_abba_full99, results='asis'}
source("~/work/hh/gompertz/full/model.R")
library(xtable)
ind99 <- grep("gompertz_abba_99_full", names(fits))
fit99 <- fits[[ind99]]
print(xtable::xtable(summary(fit99)@coef, 
                     caption = "1999 estimates",
                     display=c('s','E','E','E','E')),
                     type="html",
                     include.rownames=TRUE)
```

ABBA **2011** estimates
```{r gomp_abba_full11, results='asis'}
ind11 <- grep("gompertz_abba_11_full", names(fits))
fit11 <- fits[[ind11]]
print(xtable::xtable(summary(fit11)@coef, 
                     caption = "2011 estimates",
                     display=c('s','E','E','E','E')),
                     type="html",
                     include.rownames=TRUE)
```

-------

### ABBA figures

```{r gomp_abba_full_vis}
ps99 <- coef(fit99)
ps11 <- coef(fit11)
abbas$pred99 <- do.call("gompertz", list(ps=ps99,
                                   dbh=abbas[["DBH"]],
                                   elev=abbas[["ELEV"]],
                                   canht=abbas[["CANHT"]]))
abbas$pred11 <- do.call("gompertz", list(ps=ps11,
                                   dbh=abbas[["DBH"]],
                                   elev=abbas[["ELEV"]],
                                   canht=abbas[["CANHT"]]))
d <- melt(abbas, measure=c("HT", "pred99", "pred11"), value.name="ht", variable.name="type")
levels(d$type)[1] <- "Observed"

## Plot stuff
trans <- levels(abbas$TRAN)
for (tran in trans) {
  
  if (nrow(abbas[abbas$TRAN == tran,]) > 0) {
    
    dd <- d[d$TRAN == tran,]

    p <- ggplot(na.omit(dd), aes(DBH, ht, color=YEAR)) + 
      geom_point(data=na.omit(dd[dd$type == "Observed",]), aes(DBH, ht, color=YEAR), alpha=0.5) +
      facet_wrap(~TPLOT) + 
      geom_line(data=na.omit(dd[dd$type %in% c("pred99", "pred11"),]), aes(DBH, ht, linetype=type, color=YEAR)) +
      ggtitle(paste0("ABBA Full Gompertz model, Transect: ", tran))
    print(p)
    }
  }

```

---------------------

## BECO

**1999** estimates
```{r gomp_beco_full99, results='asis'}
source("~/work/hh/gompertz/full/model.R")
library(xtable)
ind99 <- grep("gompertz_beco_99_full", names(fits))
fit99 <- fits[[ind99]]
print(xtable::xtable(summary(fit99)@coef, 
                     caption = "1999 estimates",
                     display=c('s','E','E','E','E')),
                     type="html",
                     include.rownames=TRUE)
```

**2011** estimates
```{r gomp_beco_full11, results='asis'}
ind11 <- grep("gompertz_beco_11_full", names(fits))
fit11 <- fits[[ind11]]
print(xtable::xtable(summary(fit11)@coef, 
                     caption = "2011 estimates",
                     display=c('s','E','E','E','E')),
                     type="html",
                     include.rownames=TRUE)
```

-------

### BECO figures

```{r gomp_beco_full_vis}
ps99 <- coef(fit99)
ps11 <- coef(fit11)
becos$pred99 <- do.call("gompertz", list(ps=ps99,
                                   dbh=becos[["DBH"]],
                                   elev=becos[["ELEV"]],
                                   canht=becos[["CANHT"]]))
becos$pred11 <- do.call("gompertz", list(ps=ps11,
                                   dbh=becos[["DBH"]],
                                   elev=becos[["ELEV"]],
                                   canht=becos[["CANHT"]]))
d <- melt(becos, measure=c("HT", "pred99", "pred11"), value.name="ht", variable.name="type")
levels(d$type)[1] <- "Observed"

## Plot stuff
trans <- levels(becos$TRAN)
for (tran in trans) {
  
  if (nrow(na.omit(d[d$TRAN == tran,])) > 0) {
    
    dd <- d[d$TRAN == tran,]

    p <- ggplot(na.omit(dd), aes(DBH, ht, color=YEAR)) + 
      geom_point(data=na.omit(dd[dd$type == "Observed",]), aes(DBH, ht, color=YEAR), alpha=0.5) +
      facet_wrap(~TPLOT) + 
      geom_line(data=na.omit(dd[dd$type %in% c("pred99", "pred11"),]), aes(DBH, ht, linetype=type, color=YEAR)) +
      ggtitle(paste0("BECO Full Gompertz model, Transect: ", tran))
    print(p)
    }
  }

```


---------------

# Gompertz [DBH, CANOPY]

$$
HT = \beta e^{\log(\frac{\gamma}{\beta})e^{-\alpha DBH}}
$$

where $\gamma$ is DBH height,

$\alpha = a + a_1 CANHT$,

$\beta = b + b_1 CANHT$

--------------------------------------------------

## ABBA

ABBA **1999** estimates
```{r abba_99_can, results='asis'}
mod <- "can"
source(paste0("~/work/hh/gompertz/", mod, "/model.R"))
library(xtable)
ind99 <- grep(paste0("gompertz_abba_99_", mod), names(fits))
fit99 <- fits[[ind99]]
print(xtable::xtable(summary(fit99)@coef, 
                     caption = "1999 estimates",
                     display=c('s','E','E','E','E')),
                     type="html",
                     include.rownames=TRUE)
```

ABBA **2011** estimates
```{r abba_11_can, results='asis'}
ind11 <- grep(paste0("gompertz_abba_11_", mod), names(fits))
fit11 <- fits[[ind11]]
print(xtable::xtable(summary(fit11)@coef, 
                     caption = "2011 estimates",
                     display=c('s','E','E','E','E')),
                     type="html",
                     include.rownames=TRUE)
```

-------

### ABBA figures

```{r gomp_abba_can_vis}
source(paste0("~/work/hh/gompertz/", mod, "/model.R"))
ps99 <- coef(fit99)
ps11 <- coef(fit11)
abbas$pred99 <- do.call("gompertz", list(ps=ps99,
                                   dbh=abbas[["DBH"]],
                                   canht=abbas[["CANHT"]]))
abbas$pred11 <- do.call("gompertz", list(ps=ps11,
                                   dbh=abbas[["DBH"]],
                                   canht=abbas[["CANHT"]]))
d <- melt(abbas, measure=c("HT", "pred99", "pred11"), value.name="ht", variable.name="type")
levels(d$type)[1] <- "Observed"

## Plot stuff
trans <- levels(abbas$TRAN)
for (tran in trans) {
  
  if (nrow(abbas[abbas$TRAN == tran,]) > 0) {
    
    dd <- d[d$TRAN == tran,]

    p <- ggplot(na.omit(dd), aes(DBH, ht, color=YEAR)) + 
      geom_point(data=na.omit(dd[dd$type == "Observed",]), aes(DBH, ht, color=YEAR), alpha=0.5) +
      facet_wrap(~TPLOT) + 
      geom_line(data=na.omit(dd[dd$type %in% c("pred99", "pred11"),]), aes(DBH, ht, linetype=type, color=YEAR)) +
      ggtitle(paste0("ABBA [", toupper(mod), "] Gompertz model, Transect: ", tran))
    print(p)
    }
  }

```

---------------------

## BECO

**1999** estimates
```{r, results='asis'}
mod <- "can"
source(paste0("~/work/hh/gompertz/", mod, "/model.R"))
library(xtable)
ind99 <- grep(paste0("gompertz_beco_99_", mod), names(fits))
fit99 <- fits[[ind99]]
print(xtable::xtable(summary(fit99)@coef, 
                     caption = "1999 estimates",
                     display=c('s','E','E','E','E')),
                     type="html",
                     include.rownames=TRUE)
```

**2011** estimates
```{r, results='asis'}
ind11 <- grep(paste0("gompertz_beco_11_", mod), names(fits))
fit11 <- fits[[ind11]]
print(xtable::xtable(summary(fit11)@coef, 
                     caption = "2011 estimates",
                     display=c('s','E','E','E','E')),
                     type="html",
                     include.rownames=TRUE)
```

-------

### BECO figures

```{r gomp_beco_can_vis}
ps99 <- coef(fit99)
ps11 <- coef(fit11)
becos$pred99 <- do.call("gompertz", list(ps=ps99,
                                   dbh=becos[["DBH"]],
                                   canht=becos[["CANHT"]]))
becos$pred11 <- do.call("gompertz", list(ps=ps11,
                                   dbh=becos[["DBH"]],
                                   canht=becos[["CANHT"]]))
d <- melt(becos, measure=c("HT", "pred99", "pred11"), value.name="ht", variable.name="type")
levels(d$type)[1] <- "Observed"

## Plot stuff
trans <- levels(becos$TRAN)
for (tran in trans) {
  
  if (nrow(na.omit(d[d$TRAN == tran,])) > 0) {
    
    dd <- d[d$TRAN == tran,]

    p <- ggplot(na.omit(dd), aes(DBH, ht, color=YEAR)) + 
      geom_point(data=na.omit(dd[dd$type == "Observed",]), aes(DBH, ht, color=YEAR), alpha=0.5) +
      facet_wrap(~TPLOT) + 
      geom_line(data=na.omit(dd[dd$type %in% c("pred99", "pred11"),]), aes(DBH, ht, linetype=type, color=YEAR)) +
      ggtitle(paste0("BECO [", mod, "] Gompertz model, Transect: ", tran))
    print(p)
    }
  }

```

---------------

# Gompertz [DBH, ELEV]

$$
HT = \beta e^{\log(\frac{\gamma}{\beta})e^{-\alpha DBH}}
$$

where $\gamma$ is DBH height,

$\alpha = a + a_1 ELEV$,

$\beta = b + b_1 ELEV$

--------------------------------------------------

## ABBA

ABBA **1999** estimates
```{r abba_99_elev, results='asis'}
mod <- "elev"
source(paste0("~/work/hh/gompertz/", mod, "/model.R"))
library(xtable)
ind99 <- grep(paste0("gompertz_abba_99_", mod), names(fits))
fit99 <- fits[[ind99]]
print(xtable::xtable(summary(fit99)@coef, 
                     caption = "1999 estimates",
                     display=c('s','E','E','E','E')),
                     type="html",
                     include.rownames=TRUE)
```

ABBA **2011** estimates
```{r abba_11_elev, results='asis'}
ind11 <- grep(paste0("gompertz_abba_11_", mod), names(fits))
fit11 <- fits[[ind11]]
print(xtable::xtable(summary(fit11)@coef, 
                     caption = "2011 estimates",
                     display=c('s','E','E','E','E')),
                     type="html",
                     include.rownames=TRUE)
```

-------

### ABBA figures

```{r gomp_abba_elev_vis}
mod <- "elev"
source(paste0("~/work/hh/gompertz/", mod, "/model.R"))
ps99 <- coef(fit99)
ps11 <- coef(fit11)
abbas$pred99 <- do.call("gompertz", list(ps=ps99,
                                   dbh=abbas[["DBH"]],
                                   elev=abbas[["ELEV"]]))
abbas$pred11 <- do.call("gompertz", list(ps=ps11,
                                   dbh=abbas[["DBH"]],
                                   elev=abbas[["ELEV"]]))
d <- melt(abbas, measure=c("HT", "pred99", "pred11"), value.name="ht", variable.name="type")
levels(d$type)[1] <- "Observed"

## Plot stuff
trans <- levels(abbas$TRAN)
for (tran in trans) {
  
  if (nrow(abbas[abbas$TRAN == tran,]) > 0) {
    
    dd <- d[d$TRAN == tran,]

    p <- ggplot(na.omit(dd), aes(DBH, ht, color=YEAR)) + 
      geom_point(data=na.omit(dd[dd$type == "Observed",]), aes(DBH, ht, color=YEAR), alpha=0.5) +
      facet_wrap(~TPLOT) + 
      geom_line(data=na.omit(dd[dd$type %in% c("pred99", "pred11"),]), aes(DBH, ht, linetype=type, color=YEAR)) +
      ggtitle(paste0("ABBA [", toupper(mod), "] Gompertz model, Transect: ", tran))
    print(p)
    }
  }

```

---------------------

## BECO
**NOTE**: Currently, this only has one model predicted for 1999 that is used for both years!  Need to rework the model fit for 2011.

**1999** estimates
```{r, results='asis'}
mod <- "elev"
source(paste0("~/work/hh/gompertz/", mod, "/model.R"))
library(xtable)
ind99 <- grep(paste0("gompertz_beco_99_", mod), names(fits))
fit99 <- fits[[ind99]]
print(xtable::xtable(summary(fit99)@coef, 
                     caption = "1999 estimates",
                     display=c('s','E','E','E','E')),
                     type="html",
                     include.rownames=TRUE)
```

**2011** estimates
```{r, results='asis'}
ind11 <- grep(paste0("gompertz_beco_11_", mod), names(fits))
fit11 <- fits[[ind11]]
print(xtable::xtable(summary(fit11)@coef, 
                     caption = "2011 estimates",
                     display=c('s','E','E','E','E')),
                     type="html",
                     include.rownames=TRUE)
```

-------

### BECO figures

```{r gomp_beco_elev_vis}
mod <- "elev"
source(paste0("~/work/hh/gompertz/", mod, "/model.R"))
ps99 <- coef(fit99)
ps11 <- coef(fit11)
becos$pred99 <- do.call("gompertz", list(ps=ps99,
                                   dbh=becos[["DBH"]],
                                   elev=becos[["ELEV"]]))
becos$pred11 <- do.call("gompertz", list(ps=ps11,
                                   dbh=becos[["DBH"]],
                                   elev=becos[["ELEV"]]))
d <- melt(becos, measure=c("HT", "pred99", "pred11"), value.name="ht", variable.name="type")
levels(d$type)[1] <- "Observed"

## Plot stuff
trans <- levels(becos$TRAN)
for (tran in trans) {
  
  if (nrow(na.omit(d[d$TRAN == tran,])) > 0) {
    
    dd <- d[d$TRAN == tran,]

    p <- ggplot(na.omit(dd), aes(DBH, ht, color=YEAR)) + 
      geom_point(data=na.omit(dd[dd$type == "Observed",]), aes(DBH, ht, color=YEAR), alpha=0.5) +
      facet_wrap(~TPLOT) + 
      geom_line(data=na.omit(dd[dd$type %in% c("pred99", "pred11"),]), aes(DBH, ht, linetype=type, color=YEAR)) +
      ggtitle(paste0("BECO [", toupper(mod), "] Gompertz model, Transect: ", tran))
    print(p)
    }
  }

```

---------------

# Power [DBH, CANHT, ELEV]

$$
HT = \gamma + \alpha DBH^{\beta}
$$

where $\gamma$ is DBH height,

$\alpha = a + a_1 ELEV + a_2 ELEV + a_3 CANHT*ELEV$,

$\beta = b + b_1 ELEV + b_2 CANHT + b_3 ELEV*CANHT$

--------------------------------------------------

## ABBA

ABBA **1999** estimates
```{r abba_99_pow_full, results='asis'}
mod <- "full"
source(paste0("~/work/hh/power/", mod, "/model.R"))
library(xtable)
ind99 <- grep(paste0("power_abba_99_", mod), names(fits))
fit99 <- fits[[ind99]]
print(xtable::xtable(summary(fit99)@coef, 
                     caption = "1999 estimates",
                     display=c('s','E','E','E','E')),
                     type="html",
                     include.rownames=TRUE)
```

ABBA **2011** estimates
```{r abba_11_pow_full, results='asis'}
ind11 <- grep(paste0("power_abba_11_", mod), names(fits))
fit11 <- fits[[ind11]]
print(xtable::xtable(summary(fit11)@coef, 
                     caption = "2011 estimates",
                     display=c('s','E','E','E','E')),
                     type="html",
                     include.rownames=TRUE)
```

-------

### ABBA figures

```{r pow_abba_full_vis}
mod <- "full"
source(paste0("~/work/hh/power/", mod, "/model.R"))
ps99 <- coef(fit99)
ps11 <- coef(fit11)
abbas$pred99 <- do.call("pow", list(ps=ps99,
                                   dbh=abbas[["DBH"]],
                                   canht=abbas[["CANHT"]],
                                   elev=abbas[["ELEV"]]))
abbas$pred11 <- do.call("pow", list(ps=ps11,
                                   dbh=abbas[["DBH"]],
                                   canht=abbas[["CANHT"]],
                                   elev=abbas[["ELEV"]]))
d <- melt(abbas, measure=c("HT", "pred99", "pred11"), value.name="ht", variable.name="type")
levels(d$type)[1] <- "Observed"

## Plot stuff
trans <- levels(abbas$TRAN)
for (tran in trans) {
  
  if (nrow(abbas[abbas$TRAN == tran,]) > 0) {
    
    dd <- d[d$TRAN == tran,]

    p <- ggplot(na.omit(dd), aes(DBH, ht, color=YEAR)) + 
      geom_point(data=na.omit(dd[dd$type == "Observed",]), aes(DBH, ht, color=YEAR), alpha=0.5) +
      facet_wrap(~TPLOT) + 
      geom_line(data=na.omit(dd[dd$type %in% c("pred99", "pred11"),]), aes(DBH, ht, linetype=type, color=YEAR)) +
      ggtitle(paste0("ABBA [", toupper(mod), "] Power model, Transect: ", tran))
    print(p)
    }
  }

```

---------------------

## BECO

**1999** estimates
```{r, results='asis'}
mod <- "full"
source(paste0("~/work/hh/power/", mod, "/model.R"))
library(xtable)
ind99 <- grep(paste0("power_beco_99_", mod), names(fits))
fit99 <- fits[[ind99]]
print(xtable::xtable(summary(fit99)@coef, 
                     caption = "1999 estimates",
                     display=c('s','E','E','E','E')),
                     type="html",
                     include.rownames=TRUE)
```

**2011** estimates
```{r, results='asis'}
ind11 <- grep(paste0("power_beco_11_", mod), names(fits))
fit11 <- fits[[ind11]]
print(xtable::xtable(summary(fit11)@coef, 
                     caption = "2011 estimates",
                     display=c('s','E','E','E','E')),
                     type="html",
                     include.rownames=TRUE)
```

-------

### BECO figures

```{r pow_beco_full_vis}
ps99 <- coef(fit99)
ps11 <- coef(fit11)
becos$pred99 <- do.call("pow", list(ps=ps99,
                                   dbh=becos[["DBH"]],
                                   canht=becos[["CANHT"]],
                                   elev=becos[["ELEV"]]))
becos$pred11 <- do.call("pow", list(ps=ps11,
                                   dbh=becos[["DBH"]],
                                   canht=becos[["CANHT"]],
                                   elev=becos[["ELEV"]]))
d <- melt(becos, measure=c("HT", "pred99", "pred11"), value.name="ht", variable.name="type")
levels(d$type)[1] <- "Observed"

## Plot stuff
trans <- levels(becos$TRAN)
for (tran in trans) {
  
  if (nrow(na.omit(d[d$TRAN == tran,])) > 0) {
    
    dd <- d[d$TRAN == tran,]

    p <- ggplot(na.omit(dd), aes(DBH, ht, color=YEAR)) + 
      geom_point(data=na.omit(dd[dd$type == "Observed",]), aes(DBH, ht, color=YEAR), alpha=0.5) +
      facet_wrap(~TPLOT) + 
      geom_line(data=na.omit(dd[dd$type %in% c("pred99", "pred11"),]), aes(DBH, ht, linetype=type, color=YEAR)) +
      ggtitle(paste0("BECO [", toupper(mod), "] Power model, Transect: ", tran))
    print(p)
    }
  }

```
